---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(cpp11)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# metal

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of metal is to provide R bindings to [Apple Metal](https://developer.apple.com/metal/), which is Apple's API for GPU programming. If you have an M1 Mac, you've got a GPU built-in! This package will let you write and compile code written in the [Metal Shading Language](https://developer.apple.com/metal/Metal-Shading-Language-Specification.pdf) (or write a package that uses the header-only Metal C++ library). This is totally experimental and really was just an excuse to work through the [simple compute function tutorial](https://developer.apple.com/documentation/metal/performing_calculations_on_a_gpu) provided by Apple.

## Installation

You can install the development version of metal from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("paleolimbot/metal")
```

This will only work on MacOS. You'll need an Intel Mac with a GPU or an M1 Mac (which has a GPU built-in). If you can run this:

```{r}
library(metal)
mtl_default_device()
```

...you're good to go!

## Example

Compile Metal Shading Langauge code:

```{r}
lib <- mtl_make_library("
  kernel void add_arrays(device const float* inA,
                         device const float* inB,
                         device float* result,
                         uint index [[thread_position_in_grid]]) {
    result[index] = inA[index] + inB[index];
  }
")

lib
```

Create some buffers and execute!

```{r}
pipeline <- mtl_compute_pipeline(lib$add_arrays)
result <- mtl_buffer(123, buffer_type = "float")
in_a <- as_mtl_floats(1:123)
in_b <- as_mtl_floats(rep(2, 123))
mtl_compute_pipeline_execute(pipeline, 123, in_a, in_b, result)
mtl_buffer_convert(result)
```

## Using the bundled metal-cpp

The [metal-cpp C++ library](https://developer.apple.com/metal/cpp/) used by this package is a header-only library that you can access by `LinkingTo: metal` in another package, use `// [[Rcpp::depends(metal)]]` in Rcpp code, or `[[cpp11::linking_to(metal)]]` in cpp11 code. You'll need to set the C++ standard to C++17 or later and possibly add `-framework Metal` to your `PKG_LIBS` if you're writing a package.

```{cpp11, cxx_std = "CXX17"}
#include <cpp11.hpp>

#define NS_PRIVATE_IMPLEMENTATION
#define CA_PRIVATE_IMPLEMENTATION
#define MTL_PRIVATE_IMPLEMENTATION
#include <Metal/Metal.hpp>

[[cpp11::linking_to(cpp11)]]
[[cpp11::linking_to(metal)]]
[[cpp11::register]]
void print_default_device() {
  MTL::Device* device = MTL::CreateSystemDefaultDevice();
  Rprintf("%s", device->description()->utf8String());
}
```

```{r}
print_default_device()
```
